Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-15 paper/main.tex
Small comment

In this project assignment, we integrate the 14 \ac{SHA}-3 candidates in
the Tahoe-\ac{LAFS} and measure the performance impact. To be able to do this,
we create Python bindings for the best suited C implementations of the SHA-3
candidates, and alter the hash function calls in the source code of Tahoe.

@@

This chapter provides background knowledge for the most important software and
technologies used in this project assignment, comprising of cryptographic hash
functions (in light of the \ac{NIST} competition), the \ac{SUPERCOP} toolkit,
Tahoe-\ac{LAFS}, and lastly Python and Cython.

@@

\section{Eliminating overhead in measurements}
The following measurements were done to eliminate the overhead introducer by
our work.

\paragraph{Cython bindings}
To eliminate as much as possible of overhead introduced by our bindings, we
timed the calls from our bindings to the underlying C-function. The overhead
should be the same for all bindings, since the Cython code is the same. The
measured overhead, calculated from BLAKE can be seen in Table XX.

\input{tables/cython_overhead}

\paragraph{Modification to Tahoe-\ac{LAFS}}
Because we write to two files after each hash calculation, this will create an
overhead in the measurements that would not normaly exist in Tahoe-\ac{LAFS}.
This overhead should be about the same for all candidates, since the number of
writes are equal. The results and calculations for this overhead, based on runs
with double \ac{SHA}-256 can be seen in Table XX.

@@

This chapter will provide comments on the results made, what the numbers
mean and why there are some inconsistencies. In addition, general
thoughts on the testing procedure and the \ac{SUPERCOP} toolkit will be
presented.

\section{Validity of \ac{SUPERCOP} results}

@@

\paragraph{Extra compiler flags}

We do not compile the candidate implementation with the exact same arguments
that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
creates \emph{position independent code}, in other words machine instruction
code that can execute properly regardless of where in memory it resides. This
will make one less register available on the x86 architecture. While we cannot
directly prove that this will affect performance of the implementations. We do
however suspect it. A person writing an optimization for \ac{SUPERCOP} will have
to balance his use of resources, among them registers, to achieve maximum speed.
The loss of one register will then give a performance penalty.

@@

\subsection{Small files}

For small files, 1B and 1KB, we can clearly see that both double \ac{SHA}-256
greatly outperforms the \ac{SHA}-3 candidate implementations. The data being
hashed for these files are dominated by lengths of 63 and 106 bytes. By
comparing with \ac{SUPERCOP} results, and especially the fact that we are using
double \ac{SHA}-256, which should result in large overhead this is unexpected.
However, we believe that the results for small files may also reflect our use of
Cython and our skill. Given large amounts of call to our bindings, while the
data processed by the actual hash functions is small and therefore does not use
a lot of time, the speed of the bindings will come in to play.

@@

\subsection{Impact of double \ac{SHA}-256}
The impact of using double \ac{SHA}-256 compared to single is surprisingly
small, even if you use the best-case estimate from Table
\ref{tbl:supercop:sha256}. For 1B, 1KB and 1MB files the data hashed is almost
exclusively 106B or less. If we assume data is hashed at the performance of 576
bytes, single \ac{SHA}-256 should outperform double \ac{SHA}-256 by about 10\%.
For 1B files this difference is eaten up by external factors (other
processes), the same goes for 1KB files. For 1MB files, there seems to be a
difference, however this difference is so small that we deem it unimportant.
For the larger files the performance penalty is even smaller, so all in all we
deem the performance penalty for Tahoe-\ac{LAFS} for using double \ac{SHA}-256
as opposed to single unimportant.

@@

\subsection{Theoretical optimized results}
We have not use the fastest available implementations for all of the
candidates, we also know that our bindings creates an overhead for all calls.
Therefore we have tried to calculate the results that could have been achieved
in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
implementations in \{SUPERCOP}. The results can be seen in appendix XX.
%TODO: Fik appendix og referanse

%Tables \ref{tbl:tahoe:optimized_g} and \ref{tbl:tahoe:optimized_p} exhibit
%the results calculated using the best \ac{SUPERCOP} implementation for each
%candidate from Section \ref{sec:results:supercop}, the exact sizes for each
%call to the hash function from Section \ref{sec:results:tahoedata}, and
%interpolation to estimate the time to hash a length not recorded by
%\ac{SUPERCOP}. For lengths above 4096B, the numbers for long data lengths are
%used. The results also assume only one core of the reference processor is used.

%\input{tables/tahoe_optimized_p}
%\input{tables/tahoe_optimized_g}

@@

\section{Improvement of methods}

@@

% TODO: Discuss the BLAKE run in Results when timing was done within Cython.

>>>>

In this project assignment, we integrate the fourteen \ac{SHA}-3 candidates in
the Tahoe-\ac{LAFS} and measure the performance impact. To be able to do this,
we create Python bindings for the best suited C implementations of the SHA-3
candidates, and alter the hash function calls in the source code of Tahoe.

@@

@@

\subsection{Theoretical optimized results}
These results are calculated using the best \ac{SUPERCOP} results for each
candidate from section \ref{sec:results:supercop}, the exact sizes for each
call to the hash function from section \ref{sec:results:tahoedata}, and
interpolation to estimate the time to hash a length not recorded by
\ac{SUPERCOP}. For lengths above 4096 the results for long data lanegths are
used. The results also assume only one core of the reference processor is used.

\input{tables/tahoe_optimized_p}
\input{tables/tahoe_optimized_g}

@@

% TODO: Intro

\section{Reliability of \ac{SUPERCOP} results}

@@

\paragraph{Extra compiler flags}
%TODO: Vi har andre compiler argumenter - dette må diskuteres
We do not compile the candidate implementation with the exact same arguments
that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
creates position-independant-code, in other words machine instruction code that
can execute properly regardless of where in memory it resides. This will make
one less register available on the x86 architecture. While we cannot directly
prove that this will affect performance of the implementations. We do however
suspect it. A person writing an optimization for \ac{SUPERCOP} will have to
balance his use of resources, among them registers, to achieve maximum speed.
The loss of one register will then give a performance penalty.

@@

\subsection{Small files}
For small files, 1B and 1KB, we can clearly see that both double \ac{SHA}-256 
greatly outperforms the \ac{SHA}-3 candidate implementations. The data being 
hashed for these files are dominated by lengths of 63 and 106 bytes. By 
comparing with \ac{SUPERCOP} results, and especially the fact that we are 
using double \ac{SHA}-256, which should result in large overhead this is
unexpected. However, we believe that the results for small files may also
reflect our use of Cython and our skill. Given large amounts of call to our
bindings, while the data processed by the actual hash functions is small and
therefore does not use a lot of time, the speed of the bindings will come in to
play.

@@

%FIXME: Nevn noe en plass om at vi kanskje burde timet kallene til HASH
%funksjonen også.
\subsection{Impact of double \ac{SHA}-256}
The impact of using double \ac{SHA}-256 compared to single is suprisingly
small, even if you use the best-case estimate from Table
\ref{tbl:supercop:sha256}. For 1B, 1KB and 1MB files the data hashed is almost
exclusively 106B or less. If we assume data is hashed at the performance of 576
bytes, single \ac{SHA}-256 should outperform double \ac{SHA}-256 by about 10\%.
For 1B files this difference is eaten up by external factors (other
processes), the same goes for 1KB files. For 1MB files, there seems to be a
difference, however this difference is so small that we deem it unimportant.
For the larger files the performance penalty is even smaller, so all in all we
deem the performance penalty for Tahoe-\ac{LAFS} for using double \ac{SHA}-256
as opposed to single unimportant.

@@

@@

\section{Improvement of methods}
% More general stuff. What could have been done differently? Criticize
% ourselves.
%
% Optimized candidates - which gain could be expected from more optimized
%   candidates?

@@
